---
title: "Linear regression - model selection and overfitting"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE, error = TRUE,
                      fig.align = "center")
```

## Main ideas

- Determine how to choose the "best" model from a set of models

- Splitting data for better predictions in the future

- Understand the basic strategy for regression analysis

## Packages

```{r packages}
library(tidyverse)
library(broom)
library(patchwork)
library(fivethirtyeight)
library(olsrr)
```

## Data

```{r load_data}
air_quality <- read.csv("data/air_quality.csv")
```

# Notes

## A quick review of multiple regression

Population model:

$$y = \beta_0 + \beta_1~x_1 + \beta_2~x_2 + \cdots + \beta_k~x_k +\epsilon$$

Sample model that we use to estimate the population model:
  
$$\hat{y} = b_0 + b_1~x_1 + b_2~x_2 + \cdots + b_k~x_k$$
 
- In multiple regression, in addition to interpreting main effects, we should 
  consider whether there is an interactive effect.
  
- When assessing model fit, the adjusted $R^2$ is preferable to the $R^2$ 
  because of the penalty assessed for including extra explanatory variables 
  that are not strong predictors of the level of the response variable.
  
To perform statistical inference, we need to ensure certain conditions are
satisfied. These are:

- Linearity: The relationship between response and predictor(s) is linear
- Independence: The residuals are independent
- Normality: The residuals are nearly normally distributed
- Equal Variance: The residuals have constant variance
  
These can be assessed using diagnostic plots that involve the residuals from
the fitted model. There are fixes if some of the assumptions are violated.
  
Another important consideration with multiple regression is multicollinearity.
If the predictors are highly correlated

## Air quality fit and diagnostics

```{r mr_1}
mr_1 <- lm(Ozone ~ Solar.R + Wind + Temp, data = airquality)
tidy(mr_1)
```

```{r mr_1_aug}
mr_1_aug <- augment(mr_1)
```

```{r mr_1_independence_check}
p1 <- ggplot(mr_1_aug, aes(x = seq(nrow(mr_1_aug)), y = .resid)) + 
  geom_point() + 
  labs(x = "Index", y = "Residual value") +
  theme_minimal()
```

```{r mr_1_var_lin}
p2 <- ggplot(mr_1_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, color = "red", lty = 2) +
  labs(x = "Predicted Price", y = "Residual value") +
  theme_minimal()
```

```{r mr_1_normality_1}
p3 <- ggplot(mr_1_aug, mapping = aes(x = .resid)) +
  geom_histogram(binwidth = 10, fill = "pink", color = "grey90") + 
  labs(x = "Residuals", y = "Count") +
  theme_minimal()
```

```{r mr_1_normality_2}
p4 <- ggplot(mr_1_aug, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line() +
  theme_minimal()
```

Using `patchwork` we can link together ggplot objects using `+` and `/`.

```{r all_diagnostic_plots}
(p1 + p2) / (p3 + p4)
```

Comment on what you observe with the above plots.

It looks like the normality assumption and linearity assumptions are violated
based on our diagnostic plots.

Another plot we can examine is a histogram of the response. 

```{r ozone_histogram}
ggplot(mr_1_aug, mapping = aes(x = Ozone)) +
  geom_histogram(binwidth = 10, fill = "pink", color = "grey90") + 
  labs(x = "Ozone", y = "Count") +
  theme_minimal()
```

The extremely right skewed distribution suggests that a log transformation may 
be useful to make this look more "normal".

- In R, function `log()` is by default the natural log.

```{r log_ozone_histogram}
ggplot(air_quality, mapping = aes(x = log(Ozone))) +
  geom_histogram(binwidth = 0.4, fill = "pink", color = "grey90") +
  labs(x = "log(Ozone)", y = "Count") +
  theme_minimal()
```

## Log tranformation interpretations

```{r log_mr_2}
mr_2 <- lm(log(Ozone) ~ Solar.R + Wind + Temp, data = air_quality)
tidy(mr_2)
```

$$ {log(\widehat{Ozone})} = -0.262 + 0.003~SolarR - 0.062~Wind + 0.049~Temp$$

All else held constant, for every additional degree Fahrenheit of temperature, 
the log ozone is expected to be higher, on average, by 0.049, holding all else 
constant.

#### Working with logs

- Subtraction and logs: $log(a) − log(b) = log(a / b)$

- Natural logarithm: $e^{log(x)} = x$

- We can use these identities to "undo" the log transformation

Put this back on regular scale:

$$\begin{align}0.049 &= \text{Slope}\\
0.049 &= \dfrac{log(\text{Ozone at Temp + 1}) - log(\text{Ozone at Temp})}{1}\\
0.049 &=log\left(\dfrac{\text{Ozone at Temp + 1}}{\text{Ozone at Temp}}\right)\\
e^{0.049} &=e^{log\left(\frac{\text{Ozone at Temp + 1}}{\text{Ozone at Temp}}\right)}\\
1.05 &\approx \dfrac{\text{Ozone at Temp + 1}}{\text{Ozone at Temp}}
\end{align}$$

For every one degree increase in temperature, ozone is expected to be higher,
on average, by a factor of $\text{exp}(0.049) = 1.05$. We **multiply** instead 
of add.

To do this in R

```{r regular_scale}
mr_2 %>% 
  tidy() %>% 
  mutate(estimate = exp(estimate))
```

#### Aside: what to do when y is 0

In some cases the value of the response variable might be 0. The trick to handle
this is to add a very very small number to the value of the response variable 
for these cases so that the `log` function can still be applied.

```{r log_plus_epsilon}
log(0 + 0.0000001)
```

### Recap

- Non-constant variance is one of the most common model violations, however it 
  is usually fixable by transforming the response (y) variable.

- The most common transformation when the response variable is right skewed is 
  the log transform: $log(y)$, especially useful when the response variable is 
  (extremely) right skewed.

- When using a log transformation on the response variable the interpretation of 
  the slope changes: *"For each unit increase in x, y is expected to multiply 
  by a factor of $e^{b_k}$."*

- Another useful transformation is the square root: $\sqrt{y}$, especially 
  useful when the response variable is counts.

## Model selection

- So far, you have been told which variables to include in the model. But 
  how do researchers decide which variables to include in a model?

- We may decide to include an explanatory variable in a model to examine 
  whether a significant relationship exists between the explanatory and response 
  variable. 

- We also use models for prediction. While it may seem tempting to include as 
  many variables as possible in a model, including variables that may not be 
  important can reduce the accuracy of predictions.

### Terminology

- The model with all of the variables is called the **full model**. But the 
  full model does not always give you the best prediction. 

- You may prune variables from a model in order to achieve a **parsimonious** 
  model that gives you a more accurate predictions.

### Backward elimination and forward selection

#### Backwards elimination

- Start with a **full** model (including all candidate explanatory variables 
  and all candidate interactions).

- Remove one variable at a time, and select the model with the highest 
  adjusted $R^2$.

- Continue until adjusted $R^2$ does not increase.

- A similar approach is to look at p-values. The algorithm will remove
  predictors based on a threshold level for p-values.

#### Forward selection

- Start with an **empty** model.

- Add one variable (or interaction effect) at a time, and select the model with 
  the highest adjusted $R^2$.

- Continue until adjusted $R^2$ does not increase.


#### Model selection and interaction effects

If an interaction is included in the model, the main effects of both of
those variables must also be in the model.

If a main effect is not in the model, then its interaction should not be 
in the model.
  
### Example

#### Backwards elimination

Fit our "full" model

```{r backwards_elimination_full_model}
candy_rankings <- candy_rankings %>% 
  select(-competitorname)

mr_3 <- lm(winpercent ~ ., data = candy_rankings)
```

Do backwards elimination to find our "best" model

```{r backwards_elimination}
final_model <- ols_step_backward_p(mr_3, prem = 0.10)

tidy(final_model$model)
```

#### Forward selection

We need to fit a full model first for `ols_step_forward_p()` so it knows
which variables to consider. However, the function begins with an "empty"
model when initiating the forward selection algorithm.

```{r forward_selection_fit_full_model}
mr_4 <- lm(winpercent ~ ., data = candy_rankings)
```

Do forward selection to find our "best" model

```{r forward_selection}
final_model2 <- ols_step_forward_p(mr_4, penter = 0.10)

tidy(final_model2$model)
```

## Model overfitting 

- Getting a high adjusted $R^2$ value means that you have a model that explains 
  a lot of the variation in the dependent variable. 

- However, there is such a thing as having a model that fits the data *too well.*

- When a model explains the data you have very well, but doesn't generalize well 
  to out-of-sample data you may have a problem of **overfitting**.
  
*Doing this will be optional for your project.*

### Training and testing sets

- One thing we can do to address the issue of overfitting is to split our data 
  into training and testing datasets.

- We can build our model with the training data and then see how well it 
  generalizes to the testing data.

## Strategy for regression analysis

1. Exploratory data analysis

2. Consider possible interaction effects

3. Develop one or more tentative regression models

   a. Is one or more of the models suitable for the data?
   b. If not, revise (transform and/or scale variables) and develop new models


4. Identify the most suitable model

5. Confirm assumptions are satisfied by doing a final examination of diagnostic
   plots
   
   a. If confirmed, go to step 6
   b. If not confirmed, further revise or abandon the linear model build

6. Make inferences on the basis of the regression model

7. Make predictions on the basis of the regression model
